---
title: "Stat 420 - Final Data Analysis Project Proposal"
author: "Team Name: Team Tukey, Summer 2020, Vijayakumar Sitha Mohan, Matthew Leung, Bhusan Bathani"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---


### Team Name

- **Team Tukey** 

### Team members

- **Vijayakumar Sitha Mohan**
- **Matthew Leung**
- **Bhusan Bathani**

### Tentative Title

- Bankruptcy Prediction using Binary Classification

### Desciption of the dataset

- The dataset is about bankruptcy prediction of Polish companies. The data was collected from Emerging Markets Information Service (EMIS, [Web Link]), which is a database containing information on emerging markets around the world. The bankrupt companies were analysed in the period 2000-2012, while the still operating companies were evaluated from 2007 to 2013.


- [Bankruptcy Dataset](https://www.kaggle.com/bhadaneeraj/bankruptcy-detection)


### Background information on the dataset

- Based on the collected data five classification cases were distinguished, that depends on the forecasting period:

- 1stYear the data contains financial rates from 1st year of the forecasting period and corresponding class label that indicates bankruptcy status after 5 years. The data contains 7027 instances (financial statements), 271 represents bankrupted companies, 6756 firms that did not bankrupt in the forecasting period.

- 2ndYear  the data contains financial rates from 2nd year of the forecasting period and corresponding class label that indicates bankruptcy status after 4 years. The data contains 10173 instances (financial statements), 400 represents bankrupted companies, 9773 firms that did not bankrupt in the forecasting period.

- 3rdYear  the data contains financial rates from 3rd year of the forecasting period and corresponding class label that indicates bankruptcy status after 3 years. The data contains 10503 instances (financial statements), 495 represents bankrupted companies, 10008 firms that did not bankrupt in the forecasting period.

- 4thYear the data contains financial rates from 4th year of the forecasting period and corresponding class label that indicates bankruptcy status after 2 years. The data contains 9792 instances (financial statements), 515 represents bankrupted companies, 9277 firms that did not bankrupt in the forecasting period.

- 5thYear  the data contains financial rates from 5th year of the forecasting period and corresponding class label that indicates bankruptcy status after 1 year. The data contains 5910 instances (financial statements), 410 represents bankrupted companies, 5500 firms that did not bankrupt in the forecasting period.

**Attribute Information:**

```{r echo=FALSE}
library(knitr)
df = read.csv("data/column_names.csv")
kable(df)
```



### Statement of interest

- build a model on the basis given features and target column to predict whether a bank is going to get bankrupt for a given set of features.
- Compare the result with different models and provide rationale behind the selection of the best model statistically and visually.

### Evidence of the data
```{r}
head(read.csv("data/1year.csv"))
```

