---
title: "Stat 420 - Final Data Analysis Project Proposal"
author: "Team Name: Team Tukey, Summer 2020, Vijayakumar Sitha Mohan, Matthew Leung, Bhusan Bathani"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---


### Team Name

- **Team Tukey** 

### Team members

- **Vijayakumar Sitha Mohan**
- **Matthew Leung (wmleung2)**
- **Bhusan Bathani**

### Tentative Title

- **Bank Marketing-Term Deposit Subscription Prediction using Binary Classification in R**

### Desciption of the dataset

- The data is related with direct marketing campaigns of a Portuguese banking institution. 
- The Bank marketing dataset has 41188 examples with 20 inputs and 1 output variable. 
- It has numerical as well as categorical attributes and
response attribute y denotes client subscribed to term deposit or not (yes or no). 
- The goal is to build models that can predict if client will subscribe to term deposit or not. 
- [Bank Marketing Dataset Source](http://archive.ics.uci.edu/ml/datasets/Bank+Marketing)

```{r echo=FALSE}
bank_df = read.csv("data/bank-additional-full.csv")
str(bank_df)
```

### Background information on the dataset

- The Bank-Marketing dataset is downloaded from UCI Machine Learning Repository and the same is available at http://archive.ics.uci.edu/ml/datasets/Bank+Marketing.
- There were 4 datasets in it from which bank-additional-full.csv is used that has all examples (41188) and
20 inputs ordered by date (from May 2008 to November 2010). There are 20 input variables and 1 output
variable (desired target). 
- The dataset has  customer data, socio-economic data, telemarketing data and some other data. Some attributes are numerical, and some are categorical. 
- The dataset was loaded in R Studio and checked for any missing values using is.na function and found that it didn’t have any missing values. So, we have a clean dataset.


**Attribute Information:**

```{r echo=FALSE}
library("knitr")
columns = read.csv("data/col_names.csv")
kable(columns)
```

- Output variable (desired target) is **Term Deposit** which is categorical binary variable.

  + y - if the client subscribed a term deposit
  
  + n - if the client doesn't subscribe a term deposit



### Statement of interest

- The goal is to build a model on the basis given features and target column to predict whether the client will subscribe a term deposit (variable y).
- Explore the dataset and provide data visually using plots.
- Cleanse and Standardize the data as needed
- Capture the metrics from the different models
- Finally choose the best model based on the collected metrics and provide the rationale behind the selection.


### Evidence of the data
```{r echo=FALSE}
head(bank_df)
```

##### References:
- [Moro et al., 2014] S. Moro, P. Cortez and P. Rita. A Data-Driven Approach to Predict the Success of Bank
Telemarketing. Decision Support Systems, Elsevier, 62:22-31, June 2014
