---
title: "Stat 420 - Final Data Analysis Project Proposal"
author: "Team Name: Team Tukey, Summer 2020, Vijayakumar Sitha Mohan, Matthew Leung, Bhusan Bathani"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---


### Team Name

- **Team Tukey** 

### Team members

- **Vijayakumar Sitha Mohan**
- **Matthew Leung**
- **Bhusan Bathani**

### Tentative Title

- Bankruptcy Prediction using Binary Classification

### Desciption of the dataset
- Standard and Poor’s (S&P) Capital IQ’s Compustat database, introduced in 1962, has been
extensively used in accounting research and practice due to its ease of access and wide coverage of fundamental financial and market data

- Bankruptcy is a legal status of a person or other entity that cannot repay the debts it owes to creditors. In most jurisdictions, bankruptcy is imposed by a court order, often initiated by the debtor. (Wikipedia)

- The Bankruptcy data was collected from COMPUSTAT for the years 1980 to 2000 and has 5436 observations with 13 variables.

- [Bankruptcy Dataset Source](https://www.kaggle.com/ankeetshk/bankruptcy-data-cdac)


### Background information on the dataset

This bankruptcy dataset contains13 variables, one of them was “DLRSN”- a categorical variable indicating default, the dependent variable of the prediction. Overall, the bankruptcy is about 14% of the entire sample.

**Attribute Information:**

```{r echo=FALSE}
library("knitr")
columns = read.csv("data/col_names.csv")
kable(columns)
```

- `DLRSN` stands for *Deletion Reason* which will be our predictor variable. It will contain the following codes.

```{r echo=FALSE}
codes = read.csv("data/dlrsn_codes.csv")
kable(codes)
```


```{r echo=FALSE}
bankruptcy_df = read.csv("data/bankruptcy.csv")
str(bankruptcy_df)
```


### Statement of interest

- build a model on the basis given features and target column to predict whether a bank is going to get bankrupt for a given set of features.
- Compare the result with different models and provide rationale behind the selection of the best model statistically and visually.

### Evidence of the data
```{r echo=FALSE}
head(bankruptcy_df)
```

